Individual Partisan Harm and Utility in Redistricting

### 2021-10-14 paper/paper.Rmd
merge
edits

---
title: Measuring Individual Partisan Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed. 
      The authors thank ____ for helpful comments and discussion.
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

```{r mini-harm, fig.cap="Three potential districting plans for Mininois and Minissouris, showing district partisan lean (top row) and total voter harm versus the counterfactuals in Figure 2(a)-(b) (bottom row; darker shading indicates more harm).**CK: Should there be a second scale here?**"}
include_graphics(here("paper/figures/mini_harm.pdf"))
```

@@

Partisan bias can be considered the classic measure of partisan symmetry.
The bias is the number of seats that would need to be reallocated to make the system symmetric at a chosen statewide proportion of votes, often 0.50.
It should not be surprising then that these measures differ often and greatly.

Further, by relying on alternative districting plans, we do not necessitate the use of partisan swings.
This is of great benefit in scenarios where swings must be so significant in magnitude that they stray far from the data.
Instead, the measures of harm and utility rely on $votes^{p}$ which can be model-free estimates of electoral performance of a party (namely historical election returns) or can be some sort of model, such as precinct-level election forecasts.
Of course, the votes could even be generated by notions of partisan swings, but as a choice rather than a requirement.

Notably, as the quantity of interest is not partisan symmetry, our measures do not require the existence of a seats-votes curve.
This is consequential for the study of congressional redistricting.
As a seats-votes curve is defined for a legislature, rather than a delegation, to understand the impact of single state's congressional map, you would be forced to consider how every other state drew their maps [@katz:king:rose:21a].
Thus, if this taken as the definition of a seats-votes curve, then measures of partisan symmetry lack diagnostic power to identify gerrymanders for the delegations that academics, legal scholars, and the general public are often focused on.
The individualized focus of our measures are, however, portable to congressional, state legislative, or any other form of districting and elections.

```{=html}
<!--
1.  Integrating yields a related measure of @nagle2015measures, which has also been criticized by @deford2020symmetry.
2.  Focuses on only one part of S-V curve, which may not be relevant
-->
```
## Mean-Median Difference

>>>>

---
title: Individual Partisan Harm and Utility in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed. 
      The authors thank ____ for helpful comments and discussion.
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

```{r mini-harm, fig.cap="Three potential districting plans for Mininois and Minissouris, showing district partisan lean (top row) and total voter harm versus the counterfactuals in Figure 2(a)-(b) (bottom row; darker shading indicates more harm)."}
include_graphics(here("paper/figures/mini_harm.pdf"))
```

@@

Partisan bias is probably the most widely-used measure of partisan symmetry.
The bias, denoted by $\beta(s)$ is the number of seats that would need to be reallocated to reach proportionality for a given statewide vote share $s$ (usually $\half$).
Like our proposed measures, this quantity is defined independent of electoral data and can be estimated using a uniform swing model.

In competitive states, where the baseline election has a statewide vote share near 50%, $\beta(\half)$ is highly correlated with $\E[\seats(\dem, V, Q)\mid Q=q]$ and thus $\Delta H(q)$ as well.
But in more lopsided states, $\beta(\half)$ can give unintuitive results, which, if not interpreted carefully, can lead practictioners to make incorrect conclusions about the presence and direction of a partisan gerrymander.

Fundamentally, however, $\beta(s)$ measures partisan symmetry, and not partisan harm or utility.
It should not be surprising, then, that it can and does "disagree" about the partisan effects of a particular plan.
These "disagreements" are really just differences in the target quantity of interest.

```{=html}
<!--
1.  Integrating yields a related measure of @nagle2015measures, which has also been criticized by @deford2020symmetry.
2.  Focuses on only one part of S-V curve, which may not be relevant

Notably, as the quantity of interest is not partisan symmetry, our measures do not require the existence of a seats-votes curve. This is consequential for the study of congressional redistricting. As a seats-votes curve is defined for a legislature, rather than a delegation, to understand the impact of single state's congressional map, you would be forced to consider how every other state drew their maps [@katz:king:rose:21a]. Thus, if this taken as the definition of a seats-votes curve, then measures of partisan symmetry lack diagnostic power to identify gerrymanders for the delegations that academics, legal scholars, and the general public are often focused on. The individualized focus of our measures are, however, portable to congressional, state legislative, or any other form of districting and elections.
-->
```
## Mean-Median Difference
